Let's talk about Thurstone & Co.: An information-theoretical model for comparative judgments, and its statistical translation

### 2025-01-14 paper2.qmd
section:appendix; process:write

@@

<!-- The main reason to focus on identification as a first step is that if a parameter cannot be identified using “infinite” and error free data, we certainly cannot learn anything about it using finite data.1 So, for the time being, we assume large samples and error-free measurements [@Schuessler_et_al_2023, pp. 5] -->

@@

<!-- ![](/images/png/CJ_TM_01.png){#fig-CJ_TM_01 width=23%} -->

<!-- ![](/images/png/CJ_TM_02.png){#fig-CJ_TM_02 width=32%} -->

<!-- ![](/images/png/CJ_TM_03.png){#fig-CJ_TM_03 width=45%} -->

@@

<!-- ![](/images/png/CJ_TM_05.png){#fig-CJ_TM_05 width=44%} -->

<!-- ![](/images/png/CJ_TM_06.png){#fig-CJ_TM_06 width=59%} -->

<!-- ![](/images/png/CJ_TM_07.png){#fig-CJ_TM_07 width=68%} -->

<!-- ![](/images/png/CJ_TM_08.png){#fig-CJ_TM_08 width=68%} -->

<!-- ![](/images/png/CJ_TM_09.png){#fig-CJ_TM_09 width=68%} -->

@@

<!-- ![](/images/png/CJ_TM_11.png){#fig-CJ_TM_11 width=80%} -->

@@

<!-- ![](/images/png/CJ_TM_13.png){#fig-CJ_TM_13 width=90%} -->

@@

<!-- ![](/images/png/CJ_TM_15.png){#fig-CJ_TM_15 width=90%} -->

@@

<!-- 2. What do they do next? -->
Researchers would then proceed to identify the estimands. *Identification* refers to the process of accurately computing an estimand using an estimator. An *estimator* is a method or function that transforms data into an estimate [@Neal_2020]. *Estimates* are numerical values that approximate the estimand and are derived through *estimation*, which refers to the process of integrating data with an estimator [@Everitt_et_al_2010]. The Identification-Estimation flowchart [@McElreath_2020; @Neal_2020] in @fig-IEflow provides a visual representation of the process of transitioning from estimands to estimates.

@@

<!-- 1. What does the structural approach uses? -->
The structural approach to causal inference uses SCMs and DAGs to formally and graphically represent the presumed causal structure underlying the ACE [@Pearl_2009; @Pearl_et_al_2016; @Gross_et_al_2018; @Neal_2020]. In essence, SCMs and DAGs act as *conceptual models* that guide researchers in determining which statistical models can yield valid causal inferences, assuming the depicted causal structure of the models is correct [@McElreath_2020]. Notably, every SCM has an associated DAG [@Cinelli_et_al_2020]. @fig-IEflow provides a visual representation of the role of theoretical models in the inference process.

@@

<!-- 2. What are the advantages of SCMs and DAGs? -->
SCMs and DAGs offer two key advantages for modeling causal structures. First, they enable the representation of causal relations in a non-parametric and fully interactive manner. This feature allows for feasible ACE identification strategies without specifying the data type or the nature of the functional dependence among variables [@Morgan_et_al_2014]. Second, regardless of complexity, they can represent a wide range of causal structures using only five fundamental building blocks [@Neal_2020; @McElreath_2020]. This feature allows researchers to decompose complex structures, facilitating their analysis by focusing on the causal assumptions associated with each building block [@McElreath_2020].

<!-- 3. the SCMs -->
Figures [-@fig-dags_scms1], [-@fig-dags_scms2], [-@fig-dags_scms3], [-@fig-dags_scms4], and [-@fig-dags_scms5] display these five fundamental building blocks. The left panels of the figures show the formal mathematical models, represented by the SCMs, defined in terms of a set of *endogenous* variables $V=\{X_{1},X_{2},X_{3}\}$, a set of *exogenous* variables $E=\{e_{X1},e_{X2},e_{X3}\}$, and a set of functions $F=\{f_{X1},f_{X2},f_{X3}\}$ [@Pearl_2009; @Cinelli_et_al_2020; @Neal_2020]. Endogenous variables are those whose causal mechanisms a researcher chooses to model [@Neal_2020]. In contrast, exogenous variables represent *errors* or *disturbances* arising from omitted factors that the investigator chooses not to model explicitly [@Pearl_2009, pp. 27,68]. Lastly, the functions, referred to as *structural equations*, express the endogenous variables as non-parametric functions of other variables. These functions use the symbol '$:=$' to denote the asymmetrical causal dependence of the variables and the symbol '$\dsep$' to represent *d-separation*, a concept akin to (conditional) independence.

<!-- 4. The DAGs -->
The right panels of the figures display the complementary DAGs. A DAG consists of nodes connected by edges, where nodes represent random variables. The term *directed* means the edges extend from one node to another, with arrows indicating the direction of causal influence. The term *acyclic* signifies that the causal influences do not form loops, ensuring the influences do not cycle back on themselves [@McElreath_2020]. DAGs represent observed variables as solid black circles, while they use open circles for unobserved (latent) variables [@Morgan_et_al_2014]. Finally, the arrows in the graphs show the expected direction of causal influences among these variables. While DAGs often omit exogenous variables for simplicity (the *standard* representation), including these variables is advantageous (the *magnified* representation shown in the figures), as their inclusion helps to identify potential issues related to conditioning and confounding [@Cinelli_et_al_2020].

@@

<!-- 5. Implications of the basic building block of the SCMS and DAGs -->
A careful examination of the figures highlights the assumptions underlying these building blocks. Figures [-@fig-scm_bb1] and [-@fig-mdag_bb1] depict two unconnected nodes, representing a scenario where variables $X_{1}$ and $X_{3}$ are not causally related. Figures [-@fig-scm_bb2] and [-@fig-mdag_bb2] illustrate two connected nodes, representing a scenario where a *parent* node $X_{1}$ exerts a causal influence on a *child* node $X_{3}$. In this setup, $X_{3}$ is considered a *descendant* of $X_{1}$. Additionally, $X_{1}$ and $X_{3}$ are described as *adjacent* because an edge directly connects them. Figures [-@fig-scm_bb3] and [-@fig-mdag_bb3] depict a *chain*, where $X_{1}$ influences $X_{2}$, and $X_{2}$ influences $X_{3}$. In this configuration, $X_{1}$ is a parent node of $X_{2}$, which is a parent node of $X_{3}$. This structure creates a *directed path* between $X_{1}$ and $X_{3}$. Consequently, $X_{1}$ is an *ancestor* of $X_{3}$, and $X_{2}$ fully *mediates* the relationship between the two. Figures [-@fig-scm_bb4] and [-@fig-mdag_bb4] illustrate a *fork*, where variables $X_{1}$ and $X_{3}$ are both influenced by $X_{2}$. Here, $X_{2}$ is a parent node that *confounds* the relationship between $X_{1}$ and $X_{3}$. Finally, Figures [-@fig-scm_bb5] and [-@fig-mdag_bb5] show a *collider*, where variables $X_{1}$ and $X_{3}$ are concurrent causes of $X_{2}$. In this configuration, $X_{1}$ and $X_{3}$ are not causally related to each other but both influence $X_{2}$ (an *inmorality*). Notably, in all SCMs, the errors are assumed to be independent of each other and from all other variables in the graph, as evidenced by the pairwise relations $e_{X1} \dsep e_{X2}$, $e_{X1} \dsep e_{X3}$, and $e_{X2} \dsep e_{X3}$.

<!-- 6. continuing with the motivating example -->
Researchers can use these building blocks to represent the scenario outlined in @sec-appendix-B. Figures [-@fig-scm_example1] and [-@fig-mdag_example1] depict the plausible causal structure for this example, assuming that the variable $X$ (socio-economic status of the school) acts as a confounder in the relationship between the teaching method $T$ and the outcome $Y$. The figures show several descendant, such as $X \rightarrow T$, $X \rightarrow Y$, and $T \rightarrow Y$, and also highlight multiple pairs of unconnected nodes, evident from the relationships $e_{T} \dsep e_{X}$, $e_{T} \dsep e_{Y}$, and $e_{X} \dsep e_{Y}$. Additional, the figures depict one fork, $X \rightarrow \{T, Y\}$, and two colliders: $\{X, e_{T}\} \rightarrow T$ and $\{X, T, e_{Y}\} \rightarrow Y$.

@@

## The probabilistic implications of DAGs {#sec-appendix-D .appendix appendix-style=plain}

<!-- 1. How do we use Bayesian inference? -->
After completing the identification process outlined in @sec-appendix-B, researchers estimate the ACE using Bayesian inference methods. Bayesian inference calculates the probability of a set of parameters $\theta$, enabling researchers to use their distributions to represent the ACE. The approach begins by defining two probability distributions: the likelihood of the data, $P(X_{1}, X_{2}, \dots, X_{n} | \theta)$, and the prior distribution, $P(\theta)$ [@Everitt_et_al_2010]. Here, $X_{n}$ represents a random variable, and for simplicity, $\theta$ denotes a parameter space of dimension one. After observing empirical data, researchers update these priors to posterior distributions using Bayes' rule [@Jeffreys_1998]:

$$
P(\theta| X_{1}, X_{2}, \dots, X_{n}) = \frac{P(X_{1}, X_{2}, \dots, X_{n} | \theta) \cdot P(\theta)}{P(X_{1}, X_{2}, \dots, X_{n})}
$$ {#eq-bayes_rule}

Researchers can further simplify the posterior updating process in @eq-bayes_rule into two steps: integrating new empirical data, which is defined by the likelihood, with the parameter update, which is determined by the priors, as demonstrated in equation @eq-prop_rule. This simplification is possible because the denominator on the right-hand side of @eq-bayes_rule acts as a normalizing constant, independent of $\theta$.

$$
P(\theta| X_{1}, X_{2}, \dots, X_{n}) \propto P(X_{1}, X_{2}, \dots, X_{n} | \theta) \cdot P(\theta)
$$ {#eq-prop_rule}

@@

<!-- 2. But what is the problem? -->
Temporarily setting aside the definition of prior distributions, it is important to note that the posterior updating process depends on the assumptions underlying the likelihood of the data. However, as the number of random variables, $n$, increases, this joint distribution becomes quickly intractable [@Neal_2020]. This is evident from @eq-like_chain, where the likelihood distribution is re-expressed in terms of chained conditional distributions.

$$
P(X_{1}, X_{2}, \dots, X_{n}|\theta) = P(X_{1}|\theta) \prod^{n}_{i=2} P(X_{i} | X_{i-1}, \dots, X_{1}, \theta )
$$ {#eq-like_chain}

<!-- 3. How do DAGs can help? -->
To address the complexity of the likelihood, researchers simplify the distribution by assuming that variables exhibit specific local (in)dependencies. These assumptions improve model tractability and streamline the estimation process. Researchers formalize these local (in)dependencies using directed acyclic graphs (DAGs). DAGs represent the probabilistic implications of these assumptions by relying on three key principles: the local Markov assumption, the minimality assumption, and the causal edges assumption. The *local Markov assumption* defines the statistical independencies implied by a DAG. It states that a variable is independent of all its non-descendants, given its parents [@Neal_2020]. The *minimality assumption* identifies the statistical dependencies implied by a DAG. It asserts that every pair of adjacent variables are dependent [@Neal_2020]. Finally, the *causal edges assumption* establishes causal relationships between parents and their children in a DAG by stating that every parent is a direct cause of their children [@Neal_2020]. Figure @fig-ACflow illustrates the flow of association and causation in graphs.

<!-- 4. What are the implications of this? -->
This lead to the *Bayesian Network factorization* of @eq-like_chain.

$$
P(X_{1}, X_{2}, \dots, X_{n}|\theta) = P(X_{1}|\theta) \prod^{n}_{i=2} P(X_{i} | pa(X_{i}), \theta )
$$ {#eq-net_factor}

<!-- Second, in a fully specified SCM, we can represent the joint distributions of n variables -->
<!-- with greater efficiency: We need only to specify the n functions that govern the relationships between the variables, and then from the probabilities of the error terms, we can discover all the probabilities that govern the joint distribution. -->

>>>>

In this section we focus on the identification of several parameters of interest from the theoretical model, together with various assessment design features relevant to CJ experiments. The main reason to focus on identification is that if these parameters cannot be identified using "infinite" and error free data, we certainly cannot learn anything about it using finite data. So, for the time being, we assume large samples and error-free measurements [@Schuessler_et_al_2023, pp. 5]

@@

@@

![](/images/png/CJ_TM_01.png){#fig-CJ_TM_01 width=23%}

![](/images/png/CJ_TM_02.png){#fig-CJ_TM_02 width=32%}

![](/images/png/CJ_TM_03.png){#fig-CJ_TM_03 width=45%}

@@

![](/images/png/CJ_TM_05.png){#fig-CJ_TM_05 width=44%}

![](/images/png/CJ_TM_06.png){#fig-CJ_TM_06 width=59%}

![](/images/png/CJ_TM_07.png){#fig-CJ_TM_07 width=68%}

![](/images/png/CJ_TM_08.png){#fig-CJ_TM_08 width=68%}

![](/images/png/CJ_TM_09.png){#fig-CJ_TM_09 width=68%}

@@

![](/images/png/CJ_TM_11.png){#fig-CJ_TM_11 width=80%}

@@

![](/images/png/CJ_TM_13.png){#fig-CJ_TM_13 width=90%}

@@

![](/images/png/CJ_TM_15.png){#fig-CJ_TM_15 width=90%}

@@

<!-- 2. What do they do next? -->
Researchers would then proceed to identify the estimands. *Identification* is concerned with determining whether an estimator can accurately compute an estimand based on its assumptions, regardless of random variability [@Schuessler_et_al_2023, pp. 4]. An *estimator* is a method or function that transforms data into an estimate [@Neal_2020]. *Estimates* are numerical values that approximate the estimand and are derived through *estimation*, which refers to the process of integrating data with an estimator [@Everitt_et_al_2010]. The Identification-Estimation flowchart [@McElreath_2020; @Neal_2020] in @fig-IEflow provides a visual representation of the process of transitioning from estimands to estimates.

@@

<!-- 1. The benefits of SCMs and DAGs -->
The structural approach to causal inference uses SCMs and DAGs to formally and graphically represent the presumed causal structure underlying the ACE [@Pearl_2009; @Pearl_et_al_2016; @Gross_et_al_2018; @Neal_2020]. Assuming the depicted causal structure is correct, SCMs and DAGs serve as *conceptual (theoretical) models* that help researchers determine which statistical models can identify an estimand, thus enabling valid causal inference [@McElreath_2020]. They achieve this by representing various causal structures using only five fundamental building blocks, regardless of complexity [@Neal_2020; @McElreath_2020]. This feature allows researchers to decompose complex structures into manageable components [@McElreath_2020]. Furthermore, SCMs and DAGs can depict causal relationships in a non-parametric and fully interactive way. This flexibility enables feasible ACE identification strategies without requiring detailed specifications of the functional relating the variables, their parameters, or their data types [@Pearl_et_al_2016]. @fig-IEflow shows how theoretical models guide the inference process.

@@

<!-- 2. the SCMs -->
Figures [-@fig-dags_scms1], [-@fig-dags_scms2], [-@fig-dags_scms3], [-@fig-dags_scms4], and [-@fig-dags_scms5] display these five fundamental building blocks. The left panels of the figures show the formal mathematical models, represented by the SCMs, defined in terms of a set of *endogenous* variables $V=\{X_{1},X_{2},X_{3}\}$, a set of *exogenous* variables $E=\{e_{X1},e_{X2},e_{X3}\}$, and a set of functions $F=\{f_{X1},f_{X2},f_{X3}\}$ [@Pearl_2009; @Cinelli_et_al_2020]. Endogenous variables are those whose causal mechanisms a researcher chooses to model [@Neal_2020]. In contrast, exogenous variables represent *errors* or *disturbances* arising from omitted factors that the investigator chooses not to model explicitly [@Pearl_2009, pp. 27,68]. Lastly, the functions, referred to as *structural equations*, express the endogenous variables as non-parametric functions of other variables. These functions use the symbol '$:=$' to denote the asymmetrical causal dependence of the variables and the symbol '$\dsep$' to represent *d-separation*, a concept akin to (conditional) independence.

<!-- 3. The DAGs -->
Notably, every SCM has an associated DAG [@Pearl_et_al_2016; @Cinelli_et_al_2020]. The right panels of the figures display these DAGs. A DAG is a graph consisting of nodes connected by edges, where the nodes represent random variables. The term *directed* means that the edges extend from one node to another, with arrows indicating the direction of causal influence. The term *acyclic* implies that the causal influences do not form loops, ensuring the influences do not cycle back on themselves [@McElreath_2020]. DAGs represent observed variables as solid black circles, while they use open circles for unobserved (latent) variables [@Morgan_et_al_2014]. Finally, the arrows in the graphs indicate the expected direction of causal influences between these variables. While DAGs often omit exogenous variables for simplicity (the *standard* representation), including these variables is advantageous (the *magnified* representation shown in the figures), as their inclusion helps to identify potential issues related to conditioning and confounding [@Cinelli_et_al_2020].

@@

<!-- 4. Theoretical implications of the basic building block of the SCMS and DAGs -->
A careful examination of the figures highlights the theoretical assumptions underlying these building blocks. Figures [-@fig-scm_bb1] and [-@fig-mdag_bb1] depict two unconnected nodes, representing a scenario where variables $X_{1}$ and $X_{3}$ are not causally related. Figures [-@fig-scm_bb2] and [-@fig-mdag_bb2] illustrate two connected nodes, representing a scenario where a *parent* node $X_{1}$ exerts a causal influence on a *child* node $X_{3}$. In this setup, $X_{3}$ is considered a *descendant* of $X_{1}$. Additionally, $X_{1}$ and $X_{3}$ are described as *adjacent* because an edge directly connects them. Figures [-@fig-scm_bb3] and [-@fig-mdag_bb3] depict a *chain*, where $X_{1}$ influences $X_{2}$, and $X_{2}$ influences $X_{3}$. In this configuration, $X_{1}$ is a parent node of $X_{2}$, which is a parent node of $X_{3}$. This structure creates a *directed path* between $X_{1}$ and $X_{3}$. Consequently, $X_{1}$ is an *ancestor* of $X_{3}$, and $X_{2}$ fully *mediates* the relationship between the two. Figures [-@fig-scm_bb4] and [-@fig-mdag_bb4] illustrate a *fork*, where variables $X_{1}$ and $X_{3}$ are both influenced by $X_{2}$. Here, $X_{2}$ is a parent node that *confounds* the relationship between $X_{1}$ and $X_{3}$. Finally, Figures [-@fig-scm_bb5] and [-@fig-mdag_bb5] show a *collider*, where variables $X_{1}$ and $X_{3}$ are concurrent causes of $X_{2}$. In this configuration, $X_{1}$ and $X_{3}$ are not causally related to each other but both influence $X_{2}$ (an *inmorality*). Notably, in all SCMs, the errors are assumed to be independent of each other and from all other variables in the graph, as evidenced by the pairwise relations $e_{X1} \dsep e_{X2}$, $e_{X1} \dsep e_{X3}$, and $e_{X2} \dsep e_{X3}$.

<!-- 5. Continuing with the motivating example -->
Researchers can use these building blocks to represent the scenario outlined in @sec-appendix-B. Figures [-@fig-scm_example1] and [-@fig-mdag_example1] depict the plausible causal structure for this example, assuming that the variable $X$ (socio-economic status of the school) acts as a confounder in the relationship between the teaching method $T$ and the outcome $Y$. The figures show several descendant, such as $X \rightarrow T$, $X \rightarrow Y$, and $T \rightarrow Y$, and also highlight multiple pairs of unconnected nodes, evident from the relationships $e_{T} \dsep e_{X}$, $e_{T} \dsep e_{Y}$, and $e_{X} \dsep e_{Y}$. Additional, the figures depict one fork, $X \rightarrow \{T, Y\}$, and two colliders: $\{X, e_{T}\} \rightarrow T$ and $\{X, T, e_{Y}\} \rightarrow Y$.

@@

## The probabilistic implications of SCMs and DAGs {#sec-appendix-D .appendix appendix-style=plain}

<!-- 1. How the SCMS and DAGs encode probabilistic information? -->
In addition to their graphical capabilities, SCMs and DAGs can encode probabilistic information about causal structures. They achieve this encoding by relying on three fundamental assumptions: the local Markov, the minimality, and the causal edges assumption. The *local Markov assumption* encodes probabilistic independencies between variables by declaring that nodes in a graph are independent of all its non-descendants, given its parents [@Neal_2020]. Meanwhile, the *minimality assumption* encodes probabilistic dependencies among variables by stating that every pair of adjacent nodes exhibits a dependency [@Neal_2020]. Finally, the *causal edges assumption* encodes causal relationships between variables by declaring that each parent node acts as a direct cause of its children [@Neal_2020]. @fig-ACflow illustrates how these assumptions influence the statistical and causal interpretations of graphs.

<!-- 2. What are the general implications of this? -->
The encoding of probabilistic information in SCMs and DAGs facilitates the efficient representation of the joint distribution of variables [@Pearl_et_al_2016, pp. 29]. A notable implication of the assumptions underlying this encoding is that any conceptual model described by an SCM and DAG can express the joint distribution of variables as a product of conditional probability distributions (CPDs) in the form $P(child \: | \: parents)$. This property is formally known as the *Bayesian Network factorization* (@eq-net_factor) [@Pearl_et_al_2016, pp. 29; @Neal_2020, pp. 21]. In this factorization, $pa(X_{i})$ denotes the set of variables that are the parents of $X_{i}$.

$$
P(X_{1}, X_{2}, \dots, X_{P}) = \prod^{P}_{p=1} P(X_{i} \: | \: pa(X_{i}) )
$$ {#eq-net_factor}

<!-- 3. What is the net result of this? -->
Ultimately, this encoding enables researchers with conceptual (theoretical) knowledge in the form of an SCM and DAG to predict patterns of (in)dependencies in the data. As highlighted by @Pearl_et_al_2016 [pp. 35], these predictions depend solely on the structure of these conceptual models without requiring the quantitative details of the equations or the distributions of the errors. Moreover, once researchers observe empirical data, the patterns of (in)dependencies in this data can provide significant insights into the validity of the proposed conceptual model.

<!-- The net result is that a researcher who has scientific knowledge in the form of a structural equation model is able to predict patterns of independencies in the data, based solely on the structure of the model’s graph, without relying on any quantitative information carried by the equations or by the distributions of the errors [@Pearl_et_al_2016, pp. 35]. -->

The concept of predicting patterns of (in)dependencies in the data is more clearly illustrated through the five fundamental building blocks outlined in @sec-appendix-C. 
<!-- For instance,the joint probability for the variables in the SCM [-@fig-scm_bb1] and DAG [-@fig-mdag_bb1] can be expressed in the following form: . Furthermore, the (in)dependencies shared by data sets generated by SCMs and DAGs with the same structure are the following: -->

<!-- SCM [-@fig-scm_bb2] and DAG [-@fig-mdag_bb2] -->

<!-- SCM [-@fig-scm_bb3] and DAG [-@fig-mdag_bb3] -->
<!-- Chain: Ux -> X -> Y -> Z <- Uz [@Pearl_et_al_2016, pp. 37] -->
<!-- 1. Z and Y are dependent -->
<!-- For some z, y, P(Z = z|Y = y) ≠ P(Z = z) -->
<!-- 2. Y and X are dependent -->
<!-- For some y, x, P(Y = y|X = x) ≠ P(Y = y) -->
<!-- 3. Z and X are likely dependent -->
<!-- For some z, x, P(Z = z|X = x) ≠ P(Z = z) -->
<!-- 4. Z and X are independent, conditional on Y -->
<!-- For all x, y, z, P(Z = z|X = x, Y = y) = P(Z = z|Y = y)' -->

<!-- SCM [-@fig-scm_bb4] and DAG [-@fig-mdag_bb4] -->
<!-- Fork: Uy -> Y <- X -> Z <- Uz [@Pearl_et_al_2016, pp. 40] -->
<!-- 1. X and Y are dependent. -->
<!-- For some x, y, P(X = x|Y = y) ≠ P(X = x) -->
<!-- 2. X and Z are dependent. -->
<!-- For some x, z, P(X = x|Z = z) ≠ P(X = x) -->
<!-- 3. Z and Y are likely dependent. -->
<!-- For some z, y, P(Z = z|Y = y) ≠ P(Z = z) -->
<!-- 4. Y and Z are independent, conditional on X. -->
<!-- For all 4x, y, z, P(Y = y|Z = z, X = x) = P(Y = y|X = x) -->

<!-- SCM [-@fig-scm_bb5] and DAG [-@fig-mdag_bb5] -->
<!-- Collider: Ux -> X -> Z <- Y <- Uy [@Pearl_et_al_2016, pp. 31] -->
<!-- 1. X and Z are dependent. -->
<!-- For some x, z, P(X = x|Z = z) ≠ P(X = x) -->
<!-- 2. Y and Z are dependent. -->
<!-- For some y, z, P(Y = y|Z = z) ≠ P(Y = y) -->
<!-- 3. X and Y are independent. -->
<!-- For all x, y, P(X = x|Y = y) = P(X = x) -->
<!-- 4. X and Y are dependent conditional on Z. -->
<!-- For some x, y, z, P(X = x|Y = y, Z = z) ≠ P(X = x|Z = z) -->

<!-- 1. How do we use Bayesian inference? -->
<!-- Researchers can estimate the ACE using Bayesian inference methods. Bayesian inference calculates the probability of a set of parameters $\theta$, enabling researchers to use their distributions to represent the ACE. The approach begins by defining two probability distributions: the likelihood of the data, $P(X_{1}, X_{2}, \dots, X_{P} | \theta)$, and the prior distribution, $P(\theta)$ [@Everitt_et_al_2010]. Here, $X_{P}$ represents a random variable, and for simplicity, $\theta$ denotes a parameter space of dimension one. After observing empirical data, researchers update these priors to posterior distributions using Bayes' rule [@Jeffreys_1998]: -->

<!-- $$ -->
<!-- P(\theta \: | \: X_{1}, X_{2}, \dots, X_{P}) = \frac{P(X_{1}, X_{2}, \dots, X_{P} \: | \: \theta) \cdot P(\theta)}{P(X_{1}, X_{2}, \dots, X_{P})} -->
<!-- $$ {#eq-bayes_rule} -->

<!-- Researchers can further simplify the posterior updating process in @eq-bayes_rule into two steps: integrating new empirical data, which is defined by the likelihood, with the parameter update, which is determined by the priors, as demonstrated in equation @eq-prop_rule. This simplification is possible because the denominator on the right-hand side of @eq-bayes_rule acts as a normalizing constant, independent of $\theta$. -->

<!-- $$ -->
<!-- P(\theta \: | \: X_{1}, X_{2}, \dots, X_{P}) \propto P(X_{1}, X_{2}, \dots, X_{P}\: | \: \theta) \cdot P(\theta) -->
<!-- $$ {#eq-prop_rule} -->

@@

<!-- 2. But what is the problem? -->
<!-- Temporarily setting aside the definition of prior distributions, it is important to note that the posterior updating process depends on the assumptions underlying the likelihood of the data. However, as the number of random variables, $p$, increases, this joint distribution becomes quickly intractable [@Neal_2020]. This is evident from @eq-like_chain, where the likelihood distribution is re-expressed in terms of chained conditional distributions.  -->

<!-- $$ -->
<!-- P(X_{1}, X_{2}, \dots, X_{P} \: | \: \theta) = P(X_{1} \: | \: \theta) \prod^{P}_{p=2} P(X_{i} \: | \: X_{i-1}, \dots, X_{1}, \theta ) -->
<!-- $$ {#eq-like_chain} -->

<!-- 3. How do DAGs can help? -->
<!-- To address the complexity of the likelihood, researchers simplify the distribution by assuming that variables exhibit specific local (in)dependencies. These assumptions improve model tractability and streamline the estimation process. Researchers formalize these local (in)dependencies using directed acyclic graphs (DAGs).  -->
